---
# Documentation: https://sourcethemes.com/academic/docs/managing-content/
title: "Forecasting With Prophet"
subtitle: ""
summary: "We forecast Wikipedia pageviews for some of the most popular singers in the free encyclopedia. We will use Facebook's Prophet, a very intuitive time series forecasting algorithm."
authors: ["Luis Dominguez"]
tags: ["r","forecast","prophet"]
categories: ["r","forecast","prophet"]
date: 2020-03-14T00:22:27-06:00
lastmod: 2020-03-14T00:22:27-06:00
featured: false
draft: false
# Focal points: Smart, Center, TopLeft, Top, TopRight, Left, Right, BottomLeft, Bottom, BottomRight.
image:
  caption: ""
  focal_point: "Smart"
  preview_only: true
projects: []
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, warning = FALSE, message = FALSE, 
                      echo = TRUE, dpi = 300, cache.lazy = FALSE,
                      tidy = "styler", fig.width = 8, fig.height = 5)
```

## Introduction

[Prophet](https://facebook.github.io/prophet/), or “Facebook Prophet,” is an open-source library for univariate (one variable) time series forecasting developed by Facebook. [Prophet](https://facebook.github.io/prophet/) implements what they refer to as an additive time series forecasting model, and the implementation supports trends, seasonality, and holidays.

## Load Data

We will use a dataset of daily pageviews of the six most popular singers on Wikipedia.

1. [Michael Jackson](https://en.wikipedia.org/wiki/Michael_Jackson)
2. [Lady Gaga](https://en.wikipedia.org/wiki/Lady_Gaga)
3. [Eminem](https://en.wikipedia.org/wiki/Eminem)
4. [Justin Bieber](https://en.wikipedia.org/wiki/Justin_Bieber)
5. [Freddie Mercury](https://en.wikipedia.org/wiki/Freddie_Mercury)
6. [Lil Wayne](https://en.wikipedia.org/wiki/Lil_Wayne)

Get the data [here.](https://pageviews.toolforge.org/?project=en.wikipedia.org&platform=all-access&agent=user&redirects=0&start=2015-07-01&end=2019-12-31&pages=Michael_Jackson|Lady_Gaga|Eminem|Justin_Bieber|Freddie_Mercury|Lil_Wayne)

```{r include=FALSE}

library(tidyverse)
library(rstan)
library(prophet)
library(lubridate)

theme_set(theme_bw() + theme(legend.position = "none"))

```


```{r data, echo=FALSE}

data <- read_csv("~/Documents/GitHub/personal_website/content/blog/forecasting-with-prophet/Top Singers Wikipedia Pageviews.csv")

```

## Data Preprocessing

Inspect the data.

```{r}
head(data)
```

This dataset has one column per singer. We have to tidy and split the data into the different singers to predict pageviews individually.

```{r}

timeseries <- data %>% 
  pivot_longer(`Michael Jackson`:`Lil Wayne`,
               names_to = "singer",
               values_to = "y") %>% 
  mutate(ds = as_date(Date)) %>% 
  select(ds,singer,y) %>% 
  arrange(singer,ds)

timeseries
```

Transforming our target variable to log pageviews will help predictions and visualization.

```{r log}

timeseries$y <- log10(timeseries$y)

```

Transform the feature singer to a factor and modifying the levels, so I can override the default alphabetical ordering.

```{r}
timeseries <- timeseries %>% 
  mutate(singer = factor(singer,levels = c("Michael Jackson" ,
                                           "Lady Gaga","Eminem",
                                           "Justin Bieber" ,  
                                           "Freddie Mercury",
                                           "Lil Wayne")))
```

We can visualize the time series facet wrapping by singer to get individual trends.

```{r}
timeseries %>% 
  ggplot(aes(ds,y,color = singer) ) +
  geom_line()+
  facet_wrap(~ singer, scales = "free_y") +
  scale_color_brewer(palette = "Set2") +
  labs(title = "Top Singers Pageviews on Wikipedia",
       y = "Log Pageviews", x = "")
  
```


## Prophet Basic Forecasting

To forecast individual singer, we are spliting their data. We will also set the last 30 days as our test dataset.

```{r}

timeseries_split <- timeseries %>% 
  split(.$singer)

timeseries_test <- map(timeseries_split,~slice_tail(.,n = 30))
timeseries_train <- map(timeseries_split,~slice_head(.,n = nrow(timeseries_split[[1]])-30))

```

We fit the model, create a future dataset, and forecast 30 days into the future.

```{r}

models <- timeseries_train %>% map(~ prophet(.))

future_datasets <- models %>% map( ~ make_future_dataframe(.,periods = 30,freq = "day"))

forecast <- map2(models,future_datasets,predict)

```


```{r}

forecast_tidy <- forecast %>% 
  do.call(rbind,.) %>% 
  mutate(singer = word(rownames(.),1,sep = "\\.")) %>% 
  left_join(timeseries, by = c("singer","ds")) %>% 
  mutate(singer = factor(singer,levels = c("Michael Jackson" ,
                                           "Lady Gaga","Eminem",
                                           "Justin Bieber" ,  
                                           "Freddie Mercury",
                                           "Lil Wayne")),
         test = ifelse(ds > (max(ds) - days(30)),1,0))

```


```{r}

forecast_tidy %>% 
  ggplot(aes(x = ds, group = singer,color = singer )) +
  geom_point(aes(y = y), size = 0.3,alpha = 0.5) +
  geom_line(aes(y = yhat),alpha=0.8, color = "black",size = 0.2) +
  geom_ribbon(aes(ymin = yhat_lower,
                  ymax = yhat_upper),color = "grey92",alpha=0.2)+
  facet_wrap(~ singer, scales = "free_y") +
  labs(title = "Top Singers Pageviews on Wikipedia",
       subtitle = "Actual obsevations in points. Forecast in lines",
       y = "Log Pageviews", x = "")


```

## Performance

In the test dataset, each singer's mean absolute error is:

```{r}

forecast_tidy %>% 
  filter(test == 1) %>% 
  mutate(diff = yhat - y) %>% 
  group_by(singer) %>% 
  summarise(mae = mean(abs(diff),na.rm = TRUE))

```

```{r}
forecast_tidy %>%
  filter(test == 1) %>%
  ggplot(aes(x = ds, group = singer,color = singer )) +
  geom_point(aes(y = y), size = 0.3,alpha = 0.5) +
  geom_line(aes(y = yhat),alpha=0.8, color = "black",size = 0.2) +
  geom_ribbon(aes(ymin = yhat_lower,
                  ymax = yhat_upper),color = "grey92",alpha=0.2)+
  facet_wrap(~ singer, scales = "free_y") +
  labs(title = "Top Singers Pageviews on Wikipedia. Test Set.",
       subtitle = "Actual obsevations in points. Forecast in lines",
       y = "Log Pageviews", x = "")
```


```{r}

forecast_tidy %>% 
  filter(test == 1) %>% 
  na.omit() %>% 
  mutate(diff = abs(yhat - y)) %>%  
  ggplot(aes(ds,diff)) +
  geom_line(color = "red",alpha = 0.5) +
  facet_wrap(~ singer) +
  labs(title = "Absolute error by day",
       y = "Abs Error of Log Pageviews")

```

The models are basic and they seem to miss the big spikes and change in trends. To improve the model, we should include big events on each of the singer's life (For example, new releases of albums or music videos).



